Structured Versus Decorated Cospans

### 2020-12-02 cospans.tex
minor edits while going over section 3

The following theorem establishes the isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.   (In \cite{BC} we took as objects finitely cocomplete categories with chosen finite colimits, just so we could turn these into symmetric monoidal categories without any further choices!!!  Should we do that here???) {\chris Can you tell if we can get away without doing that? Perhaps after you read up to before the proof?}

@@

In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting: given a monoidal category $\A$ there is a bijection between \define{monoidal opindexed categories}, which are lax monoidal pseudofunctors $F\maps\A\to\Cat$, and \define{monoidal opfibrations}, which are opfibrations $U \maps \X \to \A$ where $\X$ is monoidal, $U$ is strict monoidal, and $\otimes_\X$ preserves cocartesian liftings.   This bijection sends a monoidal opindexed category $F \maps \A \to \Cat$ to its Grothendieck construction $U \maps \inta F \to \A$.  If the monoidal structure on $\A$ is in fact cocartesian, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps\A\to\bicat{MonCat}$ where $\bicat{MonCat}$ is 2-category of monoidal categories, strong monoidal functors and monoidal natural transformations.   Further, all these correspondences lift to 2-equivalences.

@@

The first part formulates the existence of colimits \emph{locally} in each fibre, and if we let $\J$ range over all finite categories it says that the corresponding pseudofunctor $F\maps\A\to\Cat$ lands in the sub-2-category $\Rex$. The second part formulates the existence of colimits \emph{globally} in the total category $\inta F$, and if we let $\J$ range over all finite categories it says that $\X$ is finitely cocomplete and $U$ preserves all finite colimits. As an example, suppose that $\A$ has pushouts and a Grothendieck opfibration $U\colon\inta F\to\A$ has fibrewise pushouts preserved by its reindexing functors. We can construct pushouts in $\inta F$ as follows: 
\begin{equation}\label{eq:globalpushout}
 \begin{tikzcd}[ampersand replacement=\&,column sep=1in,row sep=.5in]
\& (a+_bc,w)\ar[dd,phantom, very near start,"\upback"description] \& \\
(a,x)\ar[ur] \&\& (c,z)\ar[ul] \\
\& (b,y)\ar[ul,"{\begin{cases}f\maps b\to a &\textrm{in }\A \\k\maps F(f)y\to x &\textrm{in }F(a)\end{cases}}"description]\ar[ur,"{\begin{cases}g\maps b\to c &\textrm{in }\A \\\ell\maps F(g)y\to z &\textrm{in }F(c)\end{cases}}"description] \&
 \end{tikzcd}
\end{equation}
which is the pushout $u_a\colon a\to a+_bc\leftarrow c:u_c$ in $\A$ on the first variable, and $w$ is the pushout in the fibre $F(a+_bc)$
\begin{displaymath}
\begin{tikzcd}
 & w \ar[dd,phantom, very near start,"\upback"description] & \\
 F(u_a)x\ar[ur] & & F(u_c)z\ar[ul] \\
 & F(u_a)(Ff(y))\cong F(u_c)(Fg(y))\ar[ul,"F(u_a)k"]\ar[ur,"F(u_c)\ell"'] &
 \end{tikzcd}
\end{displaymath}
In similar ways we can construct the initial object and coproducts in a Grothendieck category, namely
\begin{equation}\label{eq:fibrewisecoprod}
(a,x)+(b,y)=\left(a+ b,F(\iota_a)x+F(\iota_b)y\right), \qquad 
 0_{\inta F}=\left(0_A,0_{F(0_A)}\right)
\end{equation}
where $\iota_a:a\to a+b\leftarrow b:\iota_b$ are the sum inclusions in $\A$.

@@

Notice a subtlety in the above formulation. \cref{lem:MonGroth} provides a specific monoidal structure on the fibres which becomes \cref{eq:explicitstructure1} when the base $\A$ is cocartesian monoidal.   If this fibrewise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to any other cocartesian monoidal structure on the fibres, by the universal property of colimits. Therefore, we can simplify the above assumption to that in \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor, although we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$. For example, in that case, the coproducts induced in the Grothendieck category by the fibrewise ones as in \cref{eq:fibrewisecoprod} are equivalently given by \cref{eq:explicitstructure2}.

@@

We have now have all the necessary background to formally construct an isomorphism between the double category of decorated cospans and the double category of structured cospans, starting from a symmetric lax monoidal pseudofunctor $F \maps (\A, +) \to (\Cat, \times)$ that factors through $\Rex$ as a pseudofunctor.

@@

As a left adjoint, this induced $L$ preserves all colimits that exist between the finitely cocomplete categories $\A$ and $\inta F$, so we can construct the double category of structured cospans $_L\lCsp(\inta F)$ of \cref{thm:SC}. Its objects and vertical morphisms are those of the category $\A$ (as for $F\lCsp$ earlier), whereas horizontal 1-cells are now cospans of the form $\cspn{L(a)}{(m,x)}{L(b)}{}{}$ in the Grothendieck category $\inta F$ (\cref{def:GrothCat}). Explicitly, they consist of two pairs of morphisms
\begin{equation}\label{eq:scsphor1cell}
 (a,\bot_a)\xrightarrow{\scalebox{0.7}{\(\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}\)}}(m,x)\xleftarrow{\scalebox{0.7}{\(\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}\)}}(b,\bot_b)
\end{equation}
with $x\in F(m)$; recall that the reindexing functors $F(i),F(o)$ preserve all finite colimits across the fibre categories.
Finally, the 2-morphisms of this double category in this context unravel as follows
\begin{equation}\label{eq:2morphismLCspintaF}
 \begin{tikzcd}[row sep=1in,column sep=1.7in,ampersand replacement=\&]
 (a,\bot_a)\ar[r,"{\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}}"]\ar[d,"{\begin{cases}f\maps a\to a' &\textrm{in }\A \\\chi_a\maps F(f)(\bot_a)\cong \bot_{a'} &\textrm{in }F(a')\end{cases}}"description] \& (m,x) \ar[d,"{\begin{cases}k\maps m\to m' &\textrm{in }\A \\ \tau\maps F(k)(x)\to x' &\textrm{in }F(m')\end{cases}}"description] \& (b,\bot_b)\ar[l,"{\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}}"']\ar[d,"{\begin{cases}g\maps b\to b' &\textrm{in }\A \\\chi_b\maps F(g)(\bot_b)\cong \bot_{b'} &\textrm{in }F(b')\end{cases}}"description] \\
 (a',\bot_{a'})\ar[r,"{\begin{cases}i'\maps a'\to m' &\textrm{in }\A \\!\maps F(i')(\bot_{a'})\to x' &\textrm{in }F(m')\end{cases}}"'] \& (m',x') \& (b',\bot_{b'})\ar[l,"{\begin{cases}o'\maps b'\to m' &\textrm{in }\A \\!\maps F(o')(\bot_{b'})\to x' &\textrm{in }F(m')\end{cases}}"]
 \end{tikzcd}
\end{equation}
where the outside vertical legs come from the definition of $L$ on arrows \cref{eq:Lonarrows}. 
Explicitly, the above square commutativities translate to the conditions $k\circ i=i'\circ f$ and $k\circ o=o'\circ g$ in $\A$, and also to
\begin{gather}\label{eq:Grothcommutativity}
 F(k\circ i)(\bot_a)\cong Fk(Fi(\bot_a))\xrightarrow{Fk(!)}Fk(x)\xrightarrow{\tau}x'= \\
 F(i'\circ f)(\bot_a)\cong Fi'(Ff(\bot_a))\xrightarrow{Fi'(\chi_a)}Fi'(\bot_{a'})\xrightarrow{!}x'\nonumber
\end{gather}
in the fibre category $F(m')$, by composition in the Grothendieck category \cref{eq:compGrothcat}. Since both morphisms emanate from the mapped initial object in $F(m')$ under the finitely cocontinuous reindexing functors, they are the unique such into $x'$ therefore this gives no extra conditions; similarly for the equality including $o,o'$.

Building up to an isomorphism of the above described double categories, we now define a double functor
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp
\end{displaymath}
which on vertical categories acts as the identity $\lE_0 = \id_{\A}$. Given a horizontal 1-cell in $_L\lCsp(\inta F)$ as in \cref{eq:scsphor1cell}, $\lE_1$ maps it to the decorated cospan $i\maps a\to m\leftarrow b:o$ with decoration $x\in F(m)$,
whereas a 2-morphism of $L$-structured cospans as above is mapped to the underlying map of cospans in $\A$
\begin{displaymath}
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\end{displaymath}
in $\A$ along with the map $\tau\maps Fk(x)\to x'$ in $F(m')$, namely an $F$-decorated cospan 2-morphism. 
With these definitions, $\lE$ is indeed a (pseudo) double functor as in \cref{def:doublefun}: both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets, and there exist globular 2-isomorphisms $\lE(M) \odot \lE(N) \simrightarrow\lE(M \odot N)$, $U_{\lE m} \simrightarrow\lE(U_m)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(M)\odot \mathbb{E}(N)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b,\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn)
\end{equation}
whereas first composing $M$ and $N$ in $_L\lCsp(\inta F)$ using pushouts in $\inta F$ as in \cref{eq:globalpushout} produces 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(u_m)x+_{(\bot_{m+_bn})}F(u_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $u$ are the canonical maps into the pushout.
Under $\lE$ this is mapped to the underlying $\A$-cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(u_m)x+F(u_n)y$, since a pushout over an initial object is really a coproduct. That decoration is isomorphic to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}
F(m)\times F(n)\ar[rr,"\phi"]\ar[d,"F(u_m)\times F(u_n)"']\ar[dr,phantom,"\cong"] && F(m+n)\ar[d,"F(\psi)"]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(u_m+u_n)"description] \\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi"'] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the bottom composite is the fibrewise coproduct in $F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand side isomorphism is due to pseudonaturality of $\phi$ whereas the right-hand side isomorphism follows from universal properties and pseudofunctoriality of $F$.
Regarding identities, the identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibres. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}. %Do we want to check that E is a double functor one last time?

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we will now show that this double isomorphism is a symmetric monoidal one, namely $\lE=(\lE_0,\lE_1)$ is a symmetric strong monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0=\mathrm{id}_\A$ is a braided monoidal functor trivially. For $\lE_1$, given two structured cospans $M=(a,\bot_a)\to(m,x)\leftarrow(b,\bot_b)$ and $M'=(a',\bot_{a'})\to(m',x')\leftarrow(b',\bot_{b'})$ in $_L\lCsp(\inta F)$, we compute the appropriate decorated cospans
\begin{align*}
\lE(M)\ot\lE(M')&\;\stackrel{\cref{eq:tensordecoration}}{=}\;a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\one\xrightarrow{x\times x'}F(m)\times F(m')\xrightarrow{\phi_{m,m'}}F(m+m') \\
\lE(M\ot M')&\;\stackrel{\cref{eq:tensorstuctured}}{=}\;\lE\Big((a+a',\bot_{a+a'})\to\underbrace{(m+m',F(\iota_m)x+F(\iota_{m'})x')}_{\cref{eq:fibrewisecoprod}}\leftarrow(b+b',\bot_{b+b'})\Big)%=a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\phi_{m,m'}(x,x')
\end{align*}
which are uniquely isomorphic $F(\iota_m)x+F(\iota_{m'})\cong \phi_{m,m'}(x,x')$ in an analogous way to \cref{eq:isomorphismdiag}, as also discussed after \cref{cor:fcocMonGroth}.
Moreover, the monoidal unit for the category of arrows of $F\lCsp$ is $0_\A\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}0_\A\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}0_\A$ with the trivial decoration $\phi_0\maps \one\to F(0_A)$,
which is an initial object and as such isomorphic to $0_{F(0)}$
%=\lE((0_A,)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}(0_A,\phi_0)\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}(0_A,\phi_0))$
again by \cref{eq:explicitstructure2} and the discussion after \cref{cor:fcocMonGroth}.
It can be verified that with the above structure, $\lE_1$ is also a symmetric monoidal functor, and the rest of the axioms of \cref{defn:monoidal_double_functor} hold.
\end{proof}

>>>>

The following theorem establishes an isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.   (In \cite{BC} we took as objects finitely cocomplete categories with chosen finite colimits, just so we could turn these into symmetric monoidal categories without any further choices!!!  Should we do that here???) {\chris Can you tell if we can get away without doing that? Perhaps after you read up to before the proof?}

@@

In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting: given a monoidal category $\A$ there is a bijection between \define{monoidal opindexed categories}, which are lax monoidal pseudofunctors $F\maps\A\to\Cat$, and \define{monoidal opfibrations}, which are opfibrations $U \maps \X \to \A$ where $\X$, $\A$ are monoidal, $U$ is strict monoidal, and $\otimes_\X$ preserves cocartesian liftings.   This bijection sends a monoidal opindexed category $F \maps \A \to \Cat$ to its Grothendieck monoidal opfibration $U \maps \inta F \to \A$.  If the monoidal structure on $\A$ is in fact cocartesian, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps\A\to\bicat{MonCat}$ where $\bicat{MonCat}$ is 2-category of monoidal categories, strong monoidal functors and monoidal natural transformations. In fact, all these correspondences lift to 2-equivalences.

@@

The first part formulates the existence of colimits \emph{locally} in each fibre, and if we let $\J$ range over all finite categories it says that the corresponding pseudofunctor $F\maps\A\to\Cat$ lands in the sub-2-category $\Rex$. The second part formulates the existence of colimits \emph{globally} in the total category $\inta F$, and if we let $\J$ range over all finite categories it says that $\X$ is finitely cocomplete and $U$ preserves all finite colimits. As an example, suppose that $\A$ has pushouts and a Grothendieck opfibration $U\colon\inta F\to\A$ has fibrewise pushouts preserved by its reindexing functors. We can construct pushouts in $\inta F$ as follows: 
\begin{equation}\label{eq:globalpushout}
 \begin{tikzcd}[ampersand replacement=\&,column sep=1in,row sep=.5in]
\& (a+_bc,w)\ar[dd,phantom, very near start,"\upback"description] \& \\
(a,x)\ar[ur] \&\& (c,z)\ar[ul] \\
\& (b,y)\ar[ul,"{\begin{cases}f\maps b\to a &\textrm{in }\A \\k\maps F(f)y\to x &\textrm{in }F(a)\end{cases}}"description]\ar[ur,"{\begin{cases}g\maps b\to c &\textrm{in }\A \\\ell\maps F(g)y\to z &\textrm{in }F(c)\end{cases}}"description] \&
 \end{tikzcd}
\end{equation}
which on the first variable is given by the pushout in $\A$ on the left below, and the object $w$ is formed as the pushout in the fibre category $F(a+_bc)$ on the right below:
\begin{displaymath}
\begin{tikzcd}
& a+_bc\ar[dd,phantom, very near start,"\upback"description] & \\
a\ar[ur,"u_a"] && c\ar[ul,"u_c"'] \\
& b\ar[ul,"f"]\ar[ur,"g"'] &
\end{tikzcd}\qquad
\begin{tikzcd}
 & w \ar[dd,phantom, very near start,"\upback"description] & \\
 F(u_a)x\ar[ur] & & F(u_c)z\ar[ul] \\
 & F(u_a)(Ff(y))\cong F(u_c)(Fg(y))\ar[ul,"F(u_a)k"]\ar[ur,"F(u_c)\ell"'] &
 \end{tikzcd}
\end{displaymath}
In a similar way we can construct the initial object and coproducts in a Grothendieck category from those in the fibres, namely
\begin{equation}\label{eq:fibrewisecoprod}
(a,x)+(b,y)=\left(a+ b,F(\iota_a)x+F(\iota_b)y\right), \qquad 
 0_{\inta F}=\left(0_A,0_{F(0_A)}\right)
\end{equation}
where $\iota_a:a\to a+b\leftarrow b:\iota_b$ are the sum inclusions in $\A$ and the sum on the second variable is in the fibre $F(a+b)$.

@@

Notice a subtlety in the above formulation. \cref{lem:MonGroth} provides a specific monoidal structure on the fibres which becomes \cref{eq:explicitstructure1} when the base $\A$ is cocartesian monoidal.   If this fibrewise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to any other cocartesian monoidal structure on the fibres, by the universal property of colimits. Therefore, we can simplify the above assumption to that in \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor, although we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$. For example, in that case, the coproducts induced in the Grothendieck category from the fibrewise ones as in \cref{eq:fibrewisecoprod} are equivalently given by \cref{eq:explicitstructure2}.

@@

We now have all the necessary background to formally construct an isomorphism between the double category of decorated cospans and the double category of structured cospans, starting from a symmetric lax monoidal pseudofunctor $F \maps (\A, +) \to (\Cat, \times)$ that factors through $\Rex$ as a pseudofunctor.

@@

As a left adjoint, this induced $L$ preserves all colimits that exist between the finitely cocomplete categories $\A$ and $\inta F$, so we can construct the double category of structured cospans $_L\lCsp(\inta F)$ of \cref{thm:SC}. Its objects and vertical morphisms are those of the category $\A$ (same as for $F\lCsp$), whereas horizontal 1-cells are now cospans of the form $\cspn{L(a)}{(m,x)}{L(b)}{}{}$ in the Grothendieck category $\inta F$ (\cref{def:GrothCat}). Explicitly, they consist of two pairs of morphisms
\begin{equation}\label{eq:scsphor1cell}
 (a,\bot_a)\xrightarrow{\scalebox{0.7}{\(\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}\)}}(m,x)\xleftarrow{\scalebox{0.7}{\(\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}\)}}(b,\bot_b)
\end{equation}
with $x\in F(m)$; recall that the reindexing functors $F(i),F(o)$ preserve all finite colimits across the fibre categories.
Finally, the 2-morphisms of this double category in this context unravel as follows
\begin{equation}\label{eq:2morphismLCspintaF}
 \begin{tikzcd}[row sep=1in,column sep=1.7in,ampersand replacement=\&]
 (a,\bot_a)\ar[r,"{\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}}"]\ar[d,"{\begin{cases}f\maps a\to a' &\textrm{in }\A \\\chi_a\maps F(f)(\bot_a)\cong \bot_{a'} &\textrm{in }F(a')\end{cases}}"description] \& (m,x) \ar[d,"{\begin{cases}k\maps m\to m' &\textrm{in }\A \\ \tau\maps F(k)(x)\to x' &\textrm{in }F(m')\end{cases}}"description] \& (b,\bot_b)\ar[l,"{\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}}"']\ar[d,"{\begin{cases}g\maps b\to b' &\textrm{in }\A \\\chi_b\maps F(g)(\bot_b)\cong \bot_{b'} &\textrm{in }F(b')\end{cases}}"description] \\
 (a',\bot_{a'})\ar[r,"{\begin{cases}i'\maps a'\to m' &\textrm{in }\A \\!\maps F(i')(\bot_{a'})\to x' &\textrm{in }F(m')\end{cases}}"'] \& (m',x') \& (b',\bot_{b'})\ar[l,"{\begin{cases}o'\maps b'\to m' &\textrm{in }\A \\!\maps F(o')(\bot_{b'})\to x' &\textrm{in }F(m')\end{cases}}"]
 \end{tikzcd}
\end{equation}
where the outside vertical legs come from the definition of $L$ on arrows \cref{eq:Lonarrows}. 
Explicitly, the above square commutativities translate to the conditions $k\circ i=i'\circ f$ and $k\circ o=o'\circ g$ in $\A$, and also to
\begin{gather}\label{eq:Grothcommutativity}
 F(k\circ i)(\bot_a)\cong Fk(Fi(\bot_a))\xrightarrow{Fk(!)}Fk(x)\xrightarrow{\tau}x'= \\
 F(i'\circ f)(\bot_a)\cong Fi'(Ff(\bot_a))\xrightarrow{Fi'(\chi_a)}Fi'(\bot_{a'})\xrightarrow{!}x'\nonumber
\end{gather}
in the fibre category $F(m')$, by composition in the Grothendieck category \cref{eq:compGrothcat}. Since both morphisms emanate from the mapped initial object in $F(m')$ under the finitely cocontinuous reindexing functors, they are the unique such into $x'$ therefore this gives no extra conditions; similarly for the equality including $o,o'$.

Building up to an isomorphism of the above described double categories, we now define a double functor
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp
\end{displaymath}
which on vertical categories acts as the identity $\lE_0 = \id_{\A}$. Given a horizontal 1-cell in $_L\lCsp(\inta F)$ as in \cref{eq:scsphor1cell}, $\lE_1$ maps it to the decorated cospan $i\maps a\to m\leftarrow b:o$ with decoration $x\in F(m)$,
whereas a 2-morphism of $L$-structured cospans as in \cref{eq:2morphismLCspintaF} is mapped to the underlying map of cospans in $\A$
\begin{displaymath}
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\end{displaymath}
in $\A$ along with the map $\tau\maps Fk(x)\to x'$ in $F(m')$ coming from the middle arrow above, namely an $F$-decorated cospan 2-morphism. 
With these definitions, $\lE$ is indeed a double functor as in \cref{def:doublefun}: both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets, and there exist globular 2-isomorphisms $\lE(M) \odot \lE(N) \simrightarrow\lE(M \odot N)$, $U_{\lE m} \simrightarrow\lE(U_m)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(M)\odot \mathbb{E}(N)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b,\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn)
\end{equation}
whereas first composing $M$ and $N$ in $_L\lCsp(\inta F)$ using pushouts in $\inta F$ as in \cref{eq:globalpushout} produces 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(u_m)x+_{(\bot_{m+_bn})}F(u_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $u$ are the canonical maps into the pushout.
Under $\lE$ this is mapped to the underlying $\A$-cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(u_m)x+F(u_n)y$, since a pushout over an initial object is really a coproduct. That decoration is isomorphic to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}
F(m)\times F(n)\ar[rr,"\phi"]\ar[d,"F(u_m)\times F(u_n)"']\ar[dr,phantom,"\cong"] && F(m+n)\ar[d,"F(\psi)"]\ar[d,bend right=50, phantom,"\cong"']\ar[dl,"F(u_m+u_n)"description] \\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi"'] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the bottom composite is the fibrewise coproduct in $F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand side isomorphism is due to pseudonaturality of $\phi$ whereas the right-hand side isomorphism follows from universal properties and pseudofunctoriality of $F$.
Regarding identities, the identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibres. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}. %Do we want to check that E is a double functor one last time?

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we will now show that this double isomorphism is a symmetric monoidal one, namely $\lE=(\lE_0,\lE_1)$ is a symmetric strong monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0=\mathrm{id}_\A$ is a symmetric monoidal functor trivially. For $\lE_1$, given two structured cospans $M=(a,\bot_a)\to(m,x)\leftarrow(b,\bot_b)$ and $M'=(a',\bot_{a'})\to(m',x')\leftarrow(b',\bot_{b'})$ in $_L\lCsp(\inta F)$, we compute the appropriate decorated cospans
\begin{align*}
\lE(M)\ot\lE(M')&\;\stackrel{\cref{eq:tensordecoration}}{=}\;a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\one\xrightarrow{x\times x'}F(m)\times F(m')\xrightarrow{\phi_{m,m'}}F(m+m') \\
\lE(M\ot M')&\;\stackrel{\cref{eq:tensorstuctured}}{=}\;\lE\Big((a+a',\bot_{a+a'})\to\underbrace{(m+m',F(\iota_m)x+F(\iota_{m'})x')}_{\cref{eq:fibrewisecoprod}}\leftarrow(b+b',\bot_{b+b'})\Big)%=a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\phi_{m,m'}(x,x')
\end{align*}
which are uniquely isomorphic $F(\iota_m)x+F(\iota_{m'})\cong \phi_{m,m'}(x,x')$ in an analogous way to \cref{eq:isomorphismdiag}, as also discussed after \cref{cor:fcocMonGroth}.
Moreover, the monoidal unit for the category of arrows of $F\lCsp$ is $0_\A\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}0_\A\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}0_\A$ with the trivial decoration $\phi_0\maps \one\to F(0_A)$,
which is an initial object and as such isomorphic to $0_{F(0)}$
%=\lE((0_A,)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}(0_A,\phi_0)\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}(0_A,\phi_0))$
again by \cref{eq:explicitstructure2} and the discussion after \cref{cor:fcocMonGroth}.
It can be verified that with the above structure, $\lE_1$ is also a symmetric monoidal functor, and the rest of the axioms of \cref{defn:monoidal_double_functor} hold.
\end{proof}
